---
title: "Nucleosome Estimation"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Single-cell nucleosomes around TFBS

Below is an example clustering approach of nucleosome profiles of PBMCs around the SMC3 TF motif. Here, it was found that six distinct clusters appropriately characterize nucleosomal states around the SMC3. Different clusters exhibit varying degrees of bimodal around the binding sites, representing cells in different regulatory transition phases. 

```{r}
library(data.table)
source("/gpfs/commons/home/nmoss/rscripts/scNuc_cluster.R")

PBMC_SMC = fread("/gpfs/commons/home/nmoss/PBMC_SMC_cov_mat.txt")

ClusterNucs(coverage_matrix = PBMC_SMC, 
            clusters = 6, 
            center = FALSE)

```

We can generate heatmaps visualizing nucleosome profiles of single cells that compose each clusters. The heatmaps are ordered by NFR (nucleosomal-free region) score, a metric which assesses nucleosomal bimodality of an individual cell. 

Here is the heatmap corrosponding to cluster with the greatest proportion of cells (33.8%). Notice the strong bimodality of cells in this cluster


```{r}
PBMC_cluster5 = fread("/gpfs/commons/home/nmoss/PBMC_SMC_cluster5.txt")
HeatmapNucs(coverage_matrix = PBMC_cluster5,
            subset = T)

```

Then the heatmap of the cluster with 20.5% of cells - there is noticeable reduced definition of the +1 nucleosome. Prior to transcription-factor binding, this nucleosome is often depleted as part of an activation complex. 

```{r}
PBMC_cluster6 = fread("/gpfs/commons/home/nmoss/PBMC_SMC_cluster6.txt")
HeatmapNucs(coverage_matrix = PBMC_cluster6,
            subset = F) 
```

