---
title: "Nucleosome Estimation"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Single-cell nucleosomes around TFBS

Below is an example K-means clustering approach of nucleosome profiles of PBMCs around the SMC3 TF motif, using scATAC readouts. Per motif and tissue type, it is necessary to experiment with different numbers of clusters, due to diverse cellular differentiation states and activation mechanisms at TF-binding sites. 

At the SMC3 motif, cells exhibit varying degrees of bimodality around the motif center, often correlating with downstream expression of that TFs' target genes. 

```{r}
library(data.table)
library(RCurl)
source("https://raw.github.com/mossnicholas/scATAC/main/nucleosomes/scNuc_cluster.R")

PBMC_SMC = fread("https://raw.github.com/mossnicholas/scATAC/main/PBMC_mats/PBMC_SMC_cov_mat.txt")

ClusterNucs(coverage_matrix = PBMC_SMC, 
            clusters = 6, 
            center = FALSE)

```

We can generate heatmaps visualizing nucleosome profiles of single cells that compose each clusters. The heatmaps are ordered by NFR (nucleosomal-free region) score, a metric which assesses nucleosomal bimodality of an individual cell. 

Here are a couple heatmaps corrosponding to two of the most populated clusters above. Below, notice the strong, dense bimodal nucleosome arrangement surrounding the motif center. 


```{r}
PBMC_cluster5 = fread("https://raw.github.com/mossnicholas/scATAC/main/PBMC_mats/PBMC_SMC_cluster5.txt")
HeatmapNucs(coverage_matrix = PBMC_cluster5,
            subset = T)

```

The second example cluster has noticeablably reduced definition of the +1 nucleosome, as most cells have fewer ATAC fragments in the 100 to 200 bp range upstream. Prior to transcription-factor binding, this +1 nucleosome is often depleted as part of an activation complex. 

```{r}
PBMC_cluster6 = fread("https://raw.github.com/mossnicholas/scATAC/main/PBMC_mats/PBMC_SMC_cluster6.txt")
HeatmapNucs(coverage_matrix = PBMC_cluster6,
            subset = F) 
```

